\mytitile\\~

### 2019-10-29 document.tex

\section{Theoretical description of two-photon process}
Perturbation theory allows semianalytical calculation of the transition rates (Rabi oscillation frequencies) for single and multi-photon processes.

Let's consider the Hamiltonian $H = H_0+V(t)$, with time-dependent interaction $V(t)$ and "reference" part -- $H_0$ and wave function $\Psi(t)$. After switching to the interaction picture $V(t)$ modifies to $\tilde{V}(t) = \exp(i H_0 t)V(t)\exp(-i H_0 t)$ and Schrödinger equation gives by $i \derivative{\psi(t)}{t}=\tilde{V}(t)\psi(t)$, where $\psi(t)=\exp(i H_0 t)\Psi(t)$. The reference states of $H_0$ will be denoted $\ket{j}$, $H_0 \ket{j} = \omega_j \ket{j}$ whence it follows that matrix elements  of the modified interaction take the form $\bra{j}\tilde{V}(t)\ket{i}=\exp(i(\omega_j-\omega_i)t)\bra{j}V(t)\ket{i}$. A "solution" of Schrödinger equation in interaction picture corresponding to the initial reference state $\ket{i}$ is
\begin{equation}\label{forpsi}
	\psi_i(t)=\ket{i} - i \int_{-\infty}^{t}dt_1\tilde{V}(t_1)\ket{\psi_i(t_1)}
\end{equation}

The quantity of primary interest is not the wave function but the probability of transition from the initial state $\ket{i}\rightarrow\ket{f}$, where $\ket{f}$ is the final state of interest. Firstly, we find the transition amplitude $A_{fi} = \bra{f}U(t,t_0)\ket{i}=\bra{f}\psi_i(t)$ where $U(t,t_0)$ evolution operator.

Solving \autoref{forpsi} by simple iteration gives the series solution
\begin{equation}
	U(t,t_0) = 1+\sum_{n}U^{(n)}(t,t_0)
\end{equation}
\begin{equation}
	U^{(n)}(t,t_0) = -i^n \int_{t_0}^{t}d t_1\tilde{V}(t_1)...\int_{t_0}^{t_{n-1}}d t_n\tilde{V}(t_n)
\end{equation}
For "weak" interaction the series may be truncated to a fine term $n$ and the $n$th- order transition amplitude $\bra{f}U^{(n)}(t,t_0)\ket{i}$ can be evaluated. 
In our case the interaction in interaction picture satisfies the following relation
\begin{equation}
	\begin{split}
	\tilde{V}(t)=\exp(i H_0 t) \Omega\cos(\omega_d t)(\hat b+\hat b^{\dagger})\exp(-i H_0 t)\overset{RWA}{=}\\ \exp(i H_0 t) \Omega(\hat b^{\dagger} e^{-i\omega_d t}+\hat be^{i\omega_d t})\exp(-i H_0 t)
	\end{split}
\end{equation}
Taking the matrix element between the initial and the final states $\ket{i}$ and $\ket{f}$, one immediately obtains
\begin{equation}
	\begin{split}
	\bra{f}U^{(1)}(t,t_0)\ket{i}=-i\int_{t_0}^{t}dt_1\tilde{V}(t_1)=\\-i\frac{\Omega}{2}\big[\int_{t_0}^{t}dt_1e^{i(\omega_f-\omega_i-\omega_d)t_1}\bra{f}b^{\dagger}\ket{i}+\\
	\int_{t_0}^{t}dt_1e^{i(\omega_f-\omega_i+\omega_d)t_1}\bra{f}b\ket{i}\big]
	\end{split}
\end{equation}
To calculate second-order term we used the following expression $\sum_{j}\ket{j}\bra{j}=\mathbbm{1}$.
\begin{equation}
	\begin{split}
	\bra{f}U^{(2)}(t,t_0)\ket{i}=-\bra{f}\int_{t_0}^{t}dt_1\tilde{V}(t_1)\mathbbm{1}\int_{t_0}^{t_1}dt_2\tilde{V}(t_2)\ket{i}=\\-\int_{t_0}^{t}dt_1 e^{i(\omega_f-\omega_j)t_1}\bra{f}V(t_1)\ket{j}\int_{t_0}^{t_1}dt_2e^{i(\omega_j-\omega_i)t_2}\bra{j}V(t_2)\ket{i}
	\end{split}
\end{equation}
Go over the long-time limit $t\rightarrow\infty$ in the final integral and collect the terms belongings to the same delta functions. This gives
\begin{equation}
	\begin{split}
	\bra{f}U^{(2)}(\infty)\ket{i}=\\\frac{\pi i\Omega^2}{2}\bigg[\delta(\omega_f-\omega_i+2\omega_d)\sum_j\frac{\bra{f}b\ket{j}\bra{j}b\ket{i}}{\omega_j-\omega_i+\omega_d}+\\
	\delta(\omega_f-\omega_i)\bigg(\sum_j\frac{\bra{f}b^{\dagger}\ket{j}\bra{j}b\ket{i}}{\omega_j-\omega_i+\omega_d}+\\
	\sum_j\frac{\bra{f}b\ket{j}\bra{j}b^{\dagger}\ket{i}}{\omega_j-\omega_i-\omega_d}\bigg)+\\
	\delta(\omega_f-\omega_i-2\omega_d)\sum_j\frac{\bra{f}b^{\dagger}\ket{j}\bra{j}b^{\dagger}\ket{i}}{\omega_j-\omega_i-\omega_d}\bigg]
	\end{split}
\end{equation}

@@

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{drawing}
	\caption{Drive acts on the one of qubit, with different amplitude 100,200,300 MHz}
	\label{fig:difdrive}
\end{figure*}

 Let's described the anti-crossing highlighted in red in the picture.

>>>>

\section{Degenerate perturbation theory}
\begin{equation}
H^R=H_0+V_J
\end{equation}
At our working point the equation $\Delta_2=\alpha/2$ is true and $H_0$ becomes degenerate. Let's consider our system in the basis of dressed states.
\begin{equation}
H_0^{d}= diag(-\frac{\Omega_{1}}{2},\frac{\Omega_{1}}{2},\frac{(-\alpha-\Omega_{1})}{2},\frac{-\alpha+\Omega_{1}}{2},-\frac{\Omega_{1}}{2},\frac{\Omega_{1}}{2})
\end{equation}
The choice of wave functions from a degenerate subspace is arbitrary because any linear combination of functions from this subspace will satisfy the Schrodinger equation. However, if we demand that the change of wave functions be small under the applied small perturbation, then the choice of wave functions become defined. To obtain correct wave functions we need to diagonalize $V_J^{d}$ in the degenerate subspace. However, we can notice that $V_J^{d}$ is identically the zero matrix in our subspace, so technically any choice of basis will diagonalize $V_J^{d}$. This means that the degeneracy is not lifted at first-order in perturbation theory. We have to pick eigenvectors which diagonalize following matrix:
\begin{equation}
	\begin{matrix}
	M_{nn'} = \sum_{m}\frac{V_{nm}V_{mn'}}{E_n-E_m}
	\end{matrix}
\end{equation}
Where $\ket{n}$ and $\ket{n'}$ are states with same energy $E_n$.  The sum is over all states $m$ not in the degenerate subspace.
For example, in our case for one of degenerate subspaces:
\begin{equation}
	M_{15} =\left[\begin{matrix}\frac{J^{2} \left(\Omega_{1} - \alpha\right)}{\alpha \left(2 \Omega_{1} - \alpha\right)} & \frac{\sqrt{2} J^{2} \Omega_{1}}{\alpha \left(2 \Omega_{1} - \alpha\right)}\\\frac{\sqrt{2} J^{2} \Omega_{1}}{\alpha \left(2 \Omega_{1} - \alpha\right)} & \frac{2 J^{2} \left(\Omega_{1} - \alpha\right)}{\alpha \left(2 \Omega_{1} - \alpha\right)}\end{matrix}\right]
\end{equation}
The normalized eigenvectors $(C_n,C_{n'})$ of the matrix will give the desired zero approximation wave-functions.
The first-order correction to the states from degenerate subspace:
\begin{equation}
	\ket{n^0}+\ket{n^1} = \sum_n C_{n}\ket{n^0} + \sum_{m\ne n}\ket{m^0}\frac{\bra{m^0}V\ket{n^0}}{E_n^0-E_m^0}
\end{equation}
The sum is over all states $n$ in the degenerate subspace $n=(n,n')$.
The first-order corrections for not degenerate state compute by using the ordinary perturbation theory.
\begin{equation}
	\ket{j^1} = \sum_{m\ne j}\ket{m^0}\frac{\bra{m^0}V\ket{j^0}}{E_j^0-E_m^0}
\end{equation}

\section{Theoretical description of two-photon process}
Perturbation theory allows semianalytical calculation of the transition rates for single and multi-photon processes\cite{faisal2013theory}.

Let's consider the Hamiltonian $H = H_0+V(t)$, with time-dependent interaction $V(t)$ and "reference" part -- $H_0$ and wave function $\Psi(t)$. After switching to the interaction picture $V(t)$ modifies to $\tilde{V}(t) = \exp(i H_0 t)V(t)\exp(-i H_0 t)$ and Schrödinger equation gives by $i \derivative{\psi(t)}{t}=\tilde{V}(t)\psi(t)$, where $\psi(t)=\exp(i H_0 t)\Psi(t)$. The reference states of $H_0$ will be denoted $\ket{j}$, $H_0 \ket{j} = \omega_j \ket{j}$ whence it follows that matrix elements  of the modified interaction take the form $\bra{j}\tilde{V}(t)\ket{i}=\exp(i(\omega_j-\omega_i)t)\bra{j}V(t)\ket{i}$. A "solution" of Schrödinger equation in interaction picture corresponding to the initial reference state $\ket{i}$ is

\begin{equation}\label{forpsi}
	\tilde{\psi_i}(t)=\ket{i} - i \int_{-\infty}^{t}dt_1\tilde{V}(t_1)\ket{\tilde{\psi_i}(t_1)}
\end{equation}

Mainly important for us is to find not the wave function but the probability of transition from the initial state $\ket{i}\rightarrow\ket{f}$, where $\ket{f}$ is the final state. Firstly, we find the transition amplitude $A_{fi} = \bra{f}U(t,t_0)\ket{i}=\bra{f}\psi_i(t)$ where $U(t,t_0)$ evolution operator.

Solving \autoref{forpsi} by simple iteration gives the series solution
\begin{equation}
	U(t,t_0) = 1+\sum_{n}U^{(n)}(t,t_0)
\end{equation}
\begin{equation}
	U^{(n)}(t,t_0) = -i^n \int_{t_0}^{t}d t_1\tilde{V}(t_1)...\int_{t_0}^{t_{n-1}}d t_n\tilde{V}(t_n)
\end{equation}
For "weak" interaction the series may be truncated to a fine term $n$ and the $n$th- order transition amplitude $\bra{f}U^{(n)}(t,t_0)\ket{i}$ can be evaluated. 
In our case the interaction in interaction picture satisfies the following relation
\begin{equation}
	\begin{split}
	\tilde{V}(t)=\exp(i H_0 t) \Omega\cos(\omega_d t)(\hat b+\hat b^{\dagger})\exp(-i H_0 t)\overset{RWA}{=}\\ \exp(i H_0 t) \Omega(\hat b^{\dagger} e^{-i\omega_d t}+\hat be^{i\omega_d t})\exp(-i H_0 t)
	\end{split}
\end{equation}
Taking the matrix element between the initial and the final states $\ket{i}$ and $\ket{f}$, one immediately obtains
\begin{equation}
	\begin{split}
	\bra{f}U^{(1)}(t,t_0)\ket{i}=-i\int_{t_0}^{t}dt_1\tilde{V}(t_1)=\\-i\frac{\Omega}{2}\big[\int_{t_0}^{t}dt_1e^{i(\omega_f-\omega_i-\omega_d)t_1}\bra{f}b^{\dagger}\ket{i}+\\
	\int_{t_0}^{t}dt_1e^{i(\omega_f-\omega_i+\omega_d)t_1}\bra{f}b\ket{i}\big]
	\end{split}
\end{equation}
To calculate second-order term we used the following expression $\sum_{j}\ket{j}\bra{j}=\mathbbm{1}$.
\begin{equation}
	\begin{split}
	\bra{f}U^{(2)}(t,t_0)\ket{i}=-\bra{f}\int_{t_0}^{t}dt_1\tilde{V}(t_1)\mathbbm{1}\int_{t_0}^{t_1}dt_2\tilde{V}(t_2)\ket{i}=\\-\int_{t_0}^{t}dt_1 e^{i(\omega_f-\omega_j)t_1}\bra{f}V(t_1)\ket{j}\int_{t_0}^{t_1}dt_2e^{i(\omega_j-\omega_i)t_2}\bra{j}V(t_2)\ket{i}
	\end{split}
\end{equation}
Go over the long-time limit $t\rightarrow\infty$ in the final integral and collect the terms belongings to the same delta functions. This gives
\begin{equation}
	\begin{split}
	\bra{f}U^{(2)}(\infty)\ket{i}=\\\frac{\pi i\Omega^2}{2}\bigg[\delta(\omega_f-\omega_i+2\omega_d)\sum_j\frac{\bra{f}b\ket{j}\bra{j}b\ket{i}}{\omega_j-\omega_i+\omega_d}+\\
	\delta(\omega_f-\omega_i)\bigg(\sum_j\frac{\bra{f}b^{\dagger}\ket{j}\bra{j}b\ket{i}}{\omega_j-\omega_i+\omega_d}+\\
	\sum_j\frac{\bra{f}b\ket{j}\bra{j}b^{\dagger}\ket{i}}{\omega_j-\omega_i-\omega_d}\bigg)+\\
	\delta(\omega_f-\omega_i-2\omega_d)\sum_j\frac{\bra{f}b^{\dagger}\ket{j}\bra{j}b^{\dagger}\ket{i}}{\omega_j-\omega_i-\omega_d}\bigg]
	\end{split}
\end{equation}
The probability of two-photon transition from state $\ket{i}$ to state $\ket{f}$ is simply the square modulus of the corresponding amplitude. So, $P^{(2)}_{i\rightarrow f}= |	\bra{f}U^{(2)}(\infty)\ket{i}|^2$.
The rate (probability per unit interaction time T) of absorption of $2$ photons is defined as:
\begin{equation}
	W^{(2)}_{i\rightarrow f}=\lim\limits_{T\rightarrow\infty}\frac{P^{(2)}_{i\rightarrow f}}{T}
\end{equation}
For simplifier the equation we have made use of the relation:
\begin{equation}\nonumber
	\delta(\omega_f-\omega_i-2\omega_d) =\frac{1}{2\pi} \lim\limits_{T\rightarrow\infty}\int_{-T/2}^{T/2}e^{i(\omega_f-\omega_i-2\omega_d)t}dt = \lim\limits_{T\rightarrow\infty}\frac{T}{2\pi}
\end{equation} 
For $\omega_f=\omega_i+2\omega_d$.
\begin{equation}\label{prob}
\begin{split}
	W^{(2)}_{i\rightarrow f}=\frac{\pi\Omega^4}{8}\bigg|\sum_j\frac{\bra{f}b^{\dagger}\ket{j}\bra{j}b^{\dagger}\ket{i}}{\omega_j-\omega_i-\omega_d}\bigg|^2\delta(\omega_f-\omega_i-2\omega_d)=\\ A^{(2)}_{i\rightarrow f}\delta(\omega_f-\omega_i-2\omega_d)
\end{split}
\end{equation}

@@
